---
title: "Chapter 3 Assessment"
author: "James Farrow"
date: "06/08/2018"
output: html_document
---

```{r setup, include=FALSE}
# include any libraries you need here, you will at least need the following
library(datasets)
library(ggplot2)
# set any global chunk options using knitr::opts_chunk$set(...)
```

## Introduction

We will be looking at the `ChickWeight` dataset from the _datasets_ package.

The `ChickWeight` dataset contains `r nrow(ChickWeight)` observations of `r length(names(ChickWeight))` variables.

You can read more about `ChickWeight` by typing `?ChickWeight` in the console when _datasets_ is loaded.

Let's have a look at the first few lines of `ChickWeight`.

```{r head}
head(ChickWeight)
```

The description of `ChickWeight` is as follows:

> Weight versus age of chicks on different diets

Lets plot some aspects of `ChickWeight` to have a look.

## Discussion

First we'll plot `Time` _versus_ `weight` for the entire dataset.

```{r time-versus-weight}
ggplot(ChickWeight, aes(x=Time, y=weight)) +
  geom_point()
```

We'd expect `weight` to go up over time.
Let's add a smoothed mean to the graph using `geom_smooth()`.

```{r time-versus-weigh-with-mean}
ggplot(ChickWeight, aes(x=Time, y=weight)) +
  geom_point() +
  geom_smooth()
```

Hmm, perhaps the `Diet` might affect `weight` gain.
Let's differentiate our data points by changing the colour according to `Diet`.

```{r time-versus-weigh-with-mean-by-diet}
ggplot(ChickWeight, aes(x=Time, y=weight, color = Diet)) +
  geom_point() +
  geom_smooth()
```

That's a bit hard to read.
Facet the graphs by `Diet` as well so that each different value of `Diet` has its own plot.

```{r time-versus-weigh-with-mean-facetted-by-diet}
ggplot(ChickWeight, aes(x=Time, y=weight, color = Diet)) +
  geom_point() +
  geom_smooth() +
  facet_grid(. ~ Diet)
```

Add a layer to the graph which indicates mean ± standard error for each diet.
Draw these in black so they can be more easily seen.

```{r time-versus-weigh-with-mean-facetted-by-diet-with-mean-and-se}
ggplot(ChickWeight, aes(x=Time, y=weight, color = Diet)) +
  geom_point() +
  geom_smooth(colour = "black") +
  facet_grid(. ~ Diet)
```

Let's tidy this up a bit by adding units to the axis labels.
Add `(g)` to the y-axis label and `(days)` to the x-axis label.

```{r time-versus-weigh-with-mean-facetted-by-diet-labelled}
ggplot(ChickWeight, aes(x=Time, y=weight, color = Diet)) +
  geom_point() +
  geom_smooth(colour = "black") +
  facet_grid(. ~ Diet) +
  labs(y ="Weight (g)", x="Time (days)")
```

Finally, add a title to the plot: `Chick weight over time by diet`.

```{r time-versus-weigh-with-mean-facetted-by-diet-labelled-with-title}
ggplot(ChickWeight, aes(x=Time, y=weight, color = Diet)) +
  geom_point() +
  geom_smooth(colour = "black") +
  facet_grid(. ~ Diet) +
  labs(y ="Weight (g)", x="Time (days)", title = "Chick Weight Over Time by Diet")
```

Out of curiosity, let's try facetting by `Chick` instead (but leaving the colour set by `Diet`):

```{r time-versus-weigh-with-mean-facetted-by-chick-labelled, warning=FALSE}
ggplot(ChickWeight, aes(x=Time, y=weight, color = Diet)) +
  geom_point() +
  geom_smooth(colour = "black") +
  facet_wrap(. ~ Chick) +
  labs(y ="Weight (g)", x="Time (days)", title = "Chick Weight Over Time by Diet")
```

## Questions

  1. From the graphs, which diet seems to result in improved weight gain?
          Based on the graph, Diet 3 seems to result in the most weight gain.
  
  
  2. From the graphs, which chick(s) may have died during the course of the experiment?
          Data seemingly missing before experiment end date (the plotting of weights for some individual chicks end prior to day 20) for chicks 18, 16, 15, and 44 suggest those chicks may have died during the experiment.
